---
title: "Project Dashboard"
output: 
  flexdashboard::flex_dashboard:
    orientation: columns
    vertical_layout: fill
---

```{r setup, include=FALSE}
library(tidyverse)
library(maps)
library(mapproj)
library(usmap)
library(ggplot2)
library(flexdashboard)
library(highcharter)
library(gt)
library(htmltools)
library(viridis)
library(DT)
library(shiny)
```

```{r}
# Importing data
lungcancer_1 = read.csv("lungcancer_inc_per100k_pop_2015_2019.csv", sep = ",", header = TRUE)
lungcancer_race = read.csv("lungcancer_inc_per100k_pop_2015_2019_race.csv", sep = ",", header = TRUE)
lungcancer_gender = read.csv("lungcancer_inc_per100k_pop_2015_2019_gender.csv", sep = ",", header = TRUE)
livercancer_1 = read.csv("livercancer_inc_per100k_pop_2015_2019.csv", sep = ",", header = TRUE)
livercancer_race = read.csv("livercancer_inc_per100k_pop_2015_2019_race.csv", sep = ",", header = TRUE)
livercancer_gender = read.csv("livercancer_inc_per100k_pop_2015_2019_gender.csv", sep = ",", header = TRUE)
airquality_1 = read.csv("air_quality_pm2.5annualavg_bycounty_2018_2019.csv", sep = ",", header = TRUE)
arsenic = read.csv("arsenic_annual_mean_conc_2018_2019.csv", sep = ",", header = TRUE)
drinking = read.csv("binge_drinking_alcohol_adults_per100k_pop_2018_2019.csv", sep = ",", header = TRUE)
heart_disease = read.csv("coronary_heart_disease_per100k_pop_2018_2019.csv", sep = ",", header = TRUE)
asthma = read.csv("county_asthma_per100k_pop_2018_2019.csv", sep = ",", header = TRUE)
no_healthins = read.csv("county_noHealthIns_perc_pop_2018_2019.csv", sep = ",", header = TRUE)
diabetes = read.csv("diabetes_adults_per100k_pop_2018_2019.csv", sep = ",", header = TRUE)
obesity = read.csv("obesity_adults_per100k_pop_2018_2019.csv", sep = ",", header = TRUE)
smoking = read.csv("smoking_adults_per100k_pop_2018_2019.csv", sep = ",", header = TRUE)
poverty = read.csv("pop_perc_below_poverty_2018_2019.csv", sep = ",", header = TRUE)

# Clean up Data
livercancer_1 = subset(livercancer_1, select = -c(Data.Comment,X))
names(livercancer_1) <- tolower(names(livercancer_1))
names(livercancer_1)[names(livercancer_1) == 'countyfips'] <- 'fips'

livercancer_race = subset(livercancer_race, select = -c(Data.Comment))
names(livercancer_race)[names(livercancer_race) == 'CountyFIPS'] <- 'fips'

livercancer_racewhite <- c("White (includes Hispanic)")
livercancer_racewhite1 <- filter(livercancer_race, Race.Ethnicity %in% livercancer_racewhite)

livercancer_raceblack <- c("Black (includes Hispanic)")
livercancer_raceblack1 <- filter(livercancer_race, Race.Ethnicity %in% livercancer_raceblack)

livercancer_raceasian <- c("Asian/Pacific Islander (includes Hispanic)")
livercancer_raceasian1 <- filter(livercancer_race, Race.Ethnicity %in% livercancer_raceasian)

livercancer_raceindian <- c("American Indian/Alaskan Native (includes Hispanic)")
livercancer_raceindian1 <- filter(livercancer_race, Race.Ethnicity %in% livercancer_raceindian)

livercancer_gender = subset(livercancer_gender, select = -c(Data.Comment))
names(livercancer_gender)[names(livercancer_gender) == 'CountyFIPS'] <- 'fips'

livercancer_gendermale <- c("Male")
livercancer_gendermale1 <- filter(livercancer_gender, Gender %in% livercancer_gendermale)

livercancer_genderfemale <- c("Female")
livercancer_genderfemale1 <- filter(livercancer_gender, Gender %in% livercancer_genderfemale)

lungcancer_1 = subset(lungcancer_1, select = -c(Data.Comment,X))
names(lungcancer_1)[names(lungcancer_1) == 'CountyFIPS'] <- 'fips'

lungcancer_race = subset(lungcancer_race, select = -c(Data.Comment))
names(lungcancer_race)[names(lungcancer_race) == 'CountyFIPS'] <- 'fips'

lungcancer_racewhite <- c("White (includes Hispanic)")
lungcancer_racewhite1 <- filter(lungcancer_race, Race.Ethnicity %in% lungcancer_racewhite)

lungcancer_raceblack <- c("Black (includes Hispanic)")
lungcancer_raceblack1 <- filter(lungcancer_race, Race.Ethnicity %in% lungcancer_raceblack)

lungcancer_raceasian <- c("Asian/Pacific Islander (includes Hispanic)")
lungcancer_raceasian1 <- filter(lungcancer_race, Race.Ethnicity %in% lungcancer_raceasian)

lungcancer_raceindian <- c("American Indian/Alaskan Native (includes Hispanic)")
lungcancer_raceindian1 <- filter(lungcancer_race, Race.Ethnicity %in% livercancer_raceindian)

lungcancer_gender = subset(lungcancer_gender, select = -c(Data.Comment))
names(lungcancer_gender)[names(lungcancer_gender) == 'CountyFIPS'] <- 'fips'

lungcancer_gendermale <- c("Male")
lungcancer_gendermale1 <- filter(lungcancer_gender, Gender %in% lungcancer_gendermale)

lungcancer_genderfemale <- c("Female")
lungcancer_genderfemale1 <- filter(lungcancer_gender, Gender %in% lungcancer_genderfemale)

airquality_1 = subset(airquality_1, select = -c(Data.Comment,X))
names(airquality_1)[names(airquality_1) == 'CountyFIPS'] <- 'fips'

arsenic = subset(arsenic, select = -c(Data.Comment))
names(arsenic)[names(arsenic) == 'CountyFIPS'] <- 'fips'

drinking = subset(drinking, select = -c(Data.Comment, X))
names(drinking)[names(drinking) == 'CountyFIPS'] <- 'fips'

heart_disease = subset(heart_disease, select = -c(Data.Comment, X))
names(heart_disease)[names(heart_disease) == 'CountyFIPS'] <- 'fips'

asthma = subset(asthma, select = -c(Data.Comment, X))
names(asthma)[names(asthma) == 'CountyFIPS'] <- 'fips'

no_healthins = subset(no_healthins, select = -c(Data.Comment, X))
names(no_healthins)[names(no_healthins) == 'CountyFIPS'] <- 'fips'

diabetes = subset(diabetes, select = -c(Data.Comment, X))
names(diabetes)[names(diabetes) == 'CountyFIPS'] <- 'fips'

obesity = subset(obesity, select = -c(Data.Comment, X))
names(obesity)[names(obesity) == 'CountyFIPS'] <- 'fips'

smoking = subset(smoking, select = -c(Data.Comment, X))
names(smoking)[names(smoking) == 'CountyFIPS'] <- 'fips'

poverty = subset(poverty, select = -c(Data.Comment, X))
names(poverty)[names(poverty) == 'CountyFIPS'] <- 'fips'

```

Liver Cancer
=======================================================================

Column {.tabset .tabset-fade data-width=650}
-----------------------------------------------------------------------

### US Map

```{r}
livercancer_pt1 <- livercancer_1[, c("fips", "value")]
livercancer_pt1$value <- as.numeric(livercancer_pt1$value)
plot_usmap(regions = "counties", data = livercancer_pt1, values = "value") +
  scale_fill_gradient(low = "white", high = "red", name = "Incidence Rates") +
  labs(title = "US Counties of Liver Cancer rates") +
  theme(panel.background = element_rect(color = "black", fill = "lightblue"), legend.position = "right")
```

Column {data-width=350}
-----------------------------------------------------------------------

### Data by County

```{r}
livercancer_pt2 <- livercancer_1[, c("state", "county", "value")] 
datatable(livercancer_pt2, rownames = FALSE, options = list(pageLength =10) )
```


Liver Cancer: Race
=======================================================================

Column {.tabset .tabset-fade data-width=650}
-----------------------------------------------------------------------

### US Map Containting Race: White

```{r}
livercancer_racewhite2 <- livercancer_racewhite1[, c("fips", "Value")]
livercancer_racewhite2$Value <- as.numeric(livercancer_racewhite2$Value)
plot_usmap(regions = "counties", data = livercancer_racewhite2, values = "Value") +
  scale_fill_gradient(low = "white", high = "red", name = "Incidence Rates") +
  labs(title = "US Counties of Liver Cancer rates Of Race: White") +
  theme(panel.background = element_rect(color = "black", fill = "lightblue"), legend.position = "right")
```

### US Map Containing Race: Black

```{r}
livercancer_raceblack2 <- livercancer_raceblack1[, c("fips", "Value")]
livercancer_raceblack2$Value <- as.numeric(livercancer_raceblack2$Value)
plot_usmap(regions = "counties", data = livercancer_raceblack2, values = "Value") +
  scale_fill_gradient(low = "white", high = "red", name = "Incidence Rates") +
  labs(title = "US Counties of Liver Cancer rates Of Race: Black") +
  theme(panel.background = element_rect(color = "black", fill = "lightblue"), legend.position = "right")
```

### US Map Containing Race: Asian/Pacific Islander

```{r}
livercancer_raceasian2 <- livercancer_raceasian1[, c("fips", "Value")]
livercancer_raceasian2$Value <- as.numeric(livercancer_raceasian2$Value)
plot_usmap(regions = "counties", data = livercancer_raceasian2, values = "Value") +
  scale_fill_gradient(low = "white", high = "red", name = "Incidence Rates") +
  labs(title = "US Counties of Liver Cancer rates Of Race: Asian/Pacific Islander") +
  theme(panel.background = element_rect(color = "black", fill = "lightblue"), legend.position = "right")
```

### US Map Containing Race: American Indian/Alaskan Native

```{r}
livercancer_raceindian2 <- livercancer_raceindian1[, c("fips", "Value")]
livercancer_raceindian2$Value <- as.numeric(livercancer_raceindian2$Value)
plot_usmap(regions = "counties", data = livercancer_raceindian2, values = "Value") +
  scale_fill_gradient(low = "white", high = "red", name = "Incidence Rates") +
  labs(title = "US Counties of Liver Cancer rates Of Race: American Indian/Alaskan Native") +
  theme(panel.background = element_rect(color = "black", fill = "lightblue"), legend.position = "right")
```

Column {data-width=350}
-----------------------------------------------------------------------

### Data by Race Ethnicity

```{r}
livercancer_racept2 <- livercancer_race[, c("State", "County", "Value", "Race.Ethnicity")] 
datatable(livercancer_racept2, rownames = FALSE, options = list(pageLength =10) )
  
```

Liver Cancer: Gender
=======================================================================

Column {.tabset .tabset-fade data-width=650}
-----------------------------------------------------------------------

### US Map Containing Gender: Male

```{r}
livercancer_gendermale2 <- livercancer_gendermale1[, c("fips", "Value")]
livercancer_gendermale2$Value <- as.numeric(livercancer_gendermale2$Value)
plot_usmap(regions = "counties", data = livercancer_gendermale2, values = "Value") +
  scale_fill_gradient(low = "white", high = "red", name = "Incidence Rates") +
  labs(title = "US Counties of Liver Cancer rates Of Gender: Male") +
  theme(panel.background = element_rect(color = "black", fill = "lightblue"), legend.position = "right")
```

### US Map Containing Gender: Female

```{r}
livercancer_genderfemale2 <- livercancer_genderfemale1[, c("fips", "Value")]
livercancer_genderfemale2$Value <- as.numeric(livercancer_genderfemale2$Value)
plot_usmap(regions = "counties", data = livercancer_genderfemale2, values = "Value") +
  scale_fill_gradient(low = "white", high = "red", name = "Incidence Rates") +
  labs(title = "US Counties of Liver Cancer rates Of Gender: Female") +
  theme(panel.background = element_rect(color = "black", fill = "lightblue"), legend.position = "right")
```

Column {data-width=350}
-----------------------------------------------------------------------

### Data By Gender

```{r}
livercancer_gender2 <- livercancer_gender[, c("State", "County", "Value", "Gender")] 
datatable(livercancer_gender2, rownames = FALSE, options = list(pageLength =10) )

```


Lung Cancer
=======================================================================

Column {.tabset .tabset-fade data-width=650}
-----------------------------------------------------------------------

### US Map

```{r}
lungcancer_pt1 <- lungcancer_1[, c("fips", "Value")]
lungcancer_pt1$Value <- as.numeric(lungcancer_pt1$Value)
plot_usmap(regions = "counties", data = lungcancer_pt1, values = "Value") +
  scale_fill_gradient(low = "white", high = "green", name = "Incidence Rates Of Lung Cancer") +
  labs(title = "US Counties of Lung Cancer rates") +
  theme(panel.background = element_rect(color = "black", fill = "lightblue"), legend.position = "right")
```

Column {data-width=350}
-----------------------------------------------------------------------

### Data by County

```{r}
lungcancer_pt2 <- lungcancer_1[, c("State", "County", "Value")] 
datatable(lungcancer_pt2, rownames = FALSE, options = list(pageLength =10) )
```


Lung Cancer: Race
=======================================================================

Column {.tabset .tabset-fade data-width=650}
-----------------------------------------------------------------------

### US Map Containting Race: White

```{r}
lungcancer_racewhite2 <- lungcancer_racewhite1[, c("fips", "Value")]
lungcancer_racewhite2$Value <- as.numeric(lungcancer_racewhite2$Value)
plot_usmap(regions = "counties", data = lungcancer_racewhite2, values = "Value") +
  scale_fill_gradient(low = "white", high = "green", name = "Incidence Rates") +
  labs(title = "US Counties of Lung Cancer rates Of Race: White") +
  theme(panel.background = element_rect(color = "black", fill = "lightblue"), legend.position = "right")
```

### US Map Containing Race: Black

```{r}
lungcancer_raceblack2 <- lungcancer_raceblack1[, c("fips", "Value")]
lungcancer_raceblack2$Value <- as.numeric(lungcancer_raceblack2$Value)
plot_usmap(regions = "counties", data = lungcancer_raceblack2, values = "Value") +
  scale_fill_gradient(low = "white", high = "green", name = "Incidence Rates") +
  labs(title = "US Counties of Lung Cancer rates Of Race: Black") +
  theme(panel.background = element_rect(color = "black", fill = "lightblue"), legend.position = "right")
```

### US Map Containing Race: Asian/Pacific Islander

```{r}
lungcancer_raceasian2 <- lungcancer_raceasian1[, c("fips", "Value")]
lungcancer_raceasian2$Value <- as.numeric(lungcancer_raceasian2$Value)
plot_usmap(regions = "counties", data = lungcancer_raceasian2, values = "Value") +
  scale_fill_gradient(low = "white", high = "green", name = "Incidence Rates") +
  labs(title = "US Counties of Lung Cancer rates Of Race: Asian/Pacific Islander") +
  theme(panel.background = element_rect(color = "black", fill = "lightblue"), legend.position = "right")
```

### US Map Containing Race: American Indian/Alaskan Native

```{r}
lungcancer_raceindian2 <- lungcancer_raceindian1[, c("fips", "Value")]
lungcancer_raceindian2$Value <- as.numeric(lungcancer_raceindian2$Value)
plot_usmap(regions = "counties", data = lungcancer_raceindian2, values = "Value") +
  scale_fill_gradient(low = "white", high = "green", name = "Incidence Rates") +
  labs(title = "US Counties of Lung Cancer rates Of Race: American Indian/Alaskan Native") +
  theme(panel.background = element_rect(color = "black", fill = "lightblue"), legend.position = "right")
```

Column {data-width=350}
-----------------------------------------------------------------------

### Data by Race Ethnicity

```{r}
lungcancer_racept2 <- lungcancer_race[, c("State", "County", "Value", "Race.Ethnicity")] 
datatable(lungcancer_racept2, rownames = FALSE, options = list(pageLength =10) )
  
```

Lung Cancer: Gender
=======================================================================

Column {.tabset .tabset-fade data-width=650}
-----------------------------------------------------------------------

### US Map Containing Gender: Male

```{r}
lungcancer_gendermale2 <- lungcancer_gendermale1[, c("fips", "Value")]
lungcancer_gendermale2$Value <- as.numeric(lungcancer_gendermale2$Value)
plot_usmap(regions = "counties", data = lungcancer_gendermale2, values = "Value") +
  scale_fill_gradient(low = "white", high = "green", name = "Incidence Rates") +
  labs(title = "US Counties of Lung Cancer rates Of Gender: Male") +
  theme(panel.background = element_rect(color = "black", fill = "lightblue"), legend.position = "right")
```

### US Map Containing Gender: Female

```{r}
lungcancer_genderfemale2 <- lungcancer_genderfemale1[, c("fips", "Value")]
lungcancer_genderfemale2$Value <- as.numeric(lungcancer_genderfemale2$Value)
plot_usmap(regions = "counties", data = lungcancer_genderfemale2, values = "Value") +
  scale_fill_gradient(low = "white", high = "green", name = "Incidence Rates") +
  labs(title = "US Counties of Lung Cancer rates Of Gender: Female") +
  theme(panel.background = element_rect(color = "black", fill = "lightblue"), legend.position = "right")
```

Column {data-width=350}
-----------------------------------------------------------------------

### Data By Gender

```{r}
lungcancer_gender2 <- lungcancer_gender[, c("State", "County", "Value", "Gender")] 
datatable(lungcancer_gender2, rownames = FALSE, options = list(pageLength =10) )

```

Factors
=======================================================================

Column {.tabset .tabset-fade data-width=650}
-----------------------------------------------------------------------

### Airquality

```{r}
airquality2 <- airquality_1[, c("State", "County", "Value", "Year")] 
datatable(airquality2, rownames = FALSE, options = list(pageLength =10) )
```

### Arsenic

```{r}
arsenic1 <- arsenic[, c("State", "County", "Value", "Year")] 
datatable(arsenic1, rownames = FALSE, options = list(pageLength =10) )
```

### Drinking

```{r}
drinking1 <- drinking[, c("State", "County", "Value", "Year")] 
datatable(drinking1, rownames = FALSE, options = list(pageLength =10) )
```

### Heart Disease

```{r}
heart_disease1 <- heart_disease[, c("State", "County", "Value", "Year")] 
datatable(heart_disease1, rownames = FALSE, options = list(pageLength =10) )
```

### Asthma

```{r}
asthma1 <- asthma[, c("State", "County", "Value", "Year")] 
datatable(asthma1, rownames = FALSE, options = list(pageLength =10) )
```

### Health Insurance

```{r}
no_healthins1 <- no_healthins[, c("State", "County", "Value", "Year")] 
datatable(no_healthins1, rownames = FALSE, options = list(pageLength =10) )
```

### Diabetes

```{r}
diabetes1 <- diabetes[, c("State", "County", "Value", "Year")] 
datatable(diabetes1, rownames = FALSE, options = list(pageLength =10) )
```

### Obesity

```{r}
obesity1 <- obesity[, c("State", "County", "Value", "Year")] 
datatable(obesity1, rownames = FALSE, options = list(pageLength =10) )
```

### Smoking

```{r}
smoking1 <- smoking[, c("State", "County", "Value", "Year")] 
datatable(smoking1, rownames = FALSE, options = list(pageLength =10) )
```

### Poverty

```{r}
poverty1 <- poverty[, c("State", "County", "Value", "Year")] 
datatable(poverty1, rownames = FALSE, options = list(pageLength =10) )
```

Column {.tabset .tabset-fade data-width=350}
-----------------------------------------------------------------------

### US Map of Airquality

```{r}
airquality3 <- airquality_1[, c("fips", "Value")]
airquality3$Value <- as.numeric(airquality3$Value)
plot_usmap(regions = "counties", data = airquality3, values = "Value") +
  scale_fill_gradient(low = "white", high = "gold", name = "Incidence Rates") +
  labs(title = "US Counties Levels of Airquality") +
  theme(panel.background = element_rect(color = "black", fill = "lightblue"), legend.position = "right")
```

### US Map of Arsenic

```{r}
arsenic2 <- arsenic[, c("fips", "Value")]
arsenic2$Value <- as.numeric(arsenic2$Value)
plot_usmap(regions = "counties", data = arsenic2, values = "Value") +
  scale_fill_gradient(low = "white", high = "gold", name = "Incidence Rates") +
  labs(title = "US Counties Levels of Arsenic") +
  theme(panel.background = element_rect(color = "black", fill = "lightblue"), legend.position = "right")
```

### US Map of Drinking

```{r}
drinking2 <- drinking[, c("fips", "Value")]
drinking2$Value <- as.numeric(drinking2$Value)
plot_usmap(regions = "counties", data = drinking2, values = "Value") +
  scale_fill_gradient(low = "white", high = "gold", name = "Incidence Rates") +
  labs(title = "US Counties Levels of Drinking") +
  theme(panel.background = element_rect(color = "black", fill = "lightblue"), legend.position = "right")
```

### US Map of Heart Disease

```{r}
hd <- heart_disease[, c("fips", "Value")]
hd$Value <- as.numeric(hd$Value)
plot_usmap(regions = "counties", data = hd, values = "Value") +
  scale_fill_gradient(low = "white", high = "gold", name = "Incidence Rates") +
  labs(title = "US Counties Levels of Heart Disease") +
  theme(panel.background = element_rect(color = "black", fill = "lightblue"), legend.position = "right")
```

### US Map of Asthma

```{r}
asthma2 <- asthma[, c("fips", "Value")]
asthma2$Value <- as.numeric(asthma2$Value)
plot_usmap(regions = "counties", data = asthma2, values = "Value") +
  scale_fill_gradient(low = "white", high = "gold", name = "Incidence Rates") +
  labs(title = "US Counties Levels of Asthma") +
  theme(panel.background = element_rect(color = "black", fill = "lightblue"), legend.position = "right")
```

### US Map of Health Insurance

```{r}
hi2 <- no_healthins[, c("fips", "Value")]
hi2$Value <- as.numeric(hi2$Value)
plot_usmap(regions = "counties", data = hi2, values = "Value") +
  scale_fill_gradient(low = "white", high = "gold", name = "Incidence Rates") +
  labs(title = "US Counties With Health Insurance") +
  theme(panel.background = element_rect(color = "black", fill = "lightblue"), legend.position = "right")
```

### US Map of Diabetes

```{r}
diabetes2 <- diabetes[, c("fips", "Value")]
diabetes2$Value <- as.numeric(diabetes2$Value)
plot_usmap(regions = "counties", data = diabetes2, values = "Value") +
  scale_fill_gradient(low = "white", high = "gold", name = "Incidence Rates") +
  labs(title = "US Counties Levels of Diabetes") +
  theme(panel.background = element_rect(color = "black", fill = "lightblue"), legend.position = "right")
```

### US Map of Obesity

```{r}
obesity2 <- obesity[, c("fips", "Value")]
obesity2$Value <- as.numeric(obesity2$Value)
plot_usmap(regions = "counties", data = obesity2, values = "Value") +
  scale_fill_gradient(low = "white", high = "gold", name = "Incidence Rates") +
  labs(title = "US Counties Levels of Obesity") +
  theme(panel.background = element_rect(color = "black", fill = "lightblue"), legend.position = "right")
```

### US Map of Smoking

```{r}
smoking2 <- smoking[, c("fips", "Value")]
smoking2$Value <- as.numeric(smoking2$Value)
plot_usmap(regions = "counties", data = smoking2, values = "Value") +
  scale_fill_gradient(low = "white", high = "gold", name = "Incidence Rates") +
  labs(title = "US Counties Levels of Smoking") +
  theme(panel.background = element_rect(color = "black", fill = "lightblue"), legend.position = "right")
```


### US Map of Poverty

```{r}
poverty2 <- poverty[, c("fips", "Value")]
poverty2$Value <- as.numeric(poverty2$Value)
plot_usmap(regions = "counties", data = poverty2, values = "Value") +
  scale_fill_gradient(low = "white", high = "gold", name = "Incidence Rates") +
  labs(title = "US Counties Levels of Poverty") +
  theme(panel.background = element_rect(color = "black", fill = "lightblue"), legend.position = "right")
```

